---
title: "LNG Vessel Trip Analysis"
author: "Hannah Marszalek"
date: "2024-08-08"
output:
  html_document:
    theme: flatly
---

A note about this code: Some of the functions were written with the help of ChatGPT. While the program seems to work for the majority of the journeys, there are some that do not seem to get processed properly.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setting up

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(sf)
library(ggplot2)
library(lubridate)
library(geosphere)
library(gt)
```

### Loading in the data

```{r message=FALSE, warning=FALSE}
#ship coordinates and timestamps
all_2017 <- read_csv("data/all_2017.csv") %>% 
  select(IMO, LAT, LON, TIMESTAMP) %>% 
  distinct()

all_23_24 <- read_excel("data/lng_vessels_23_24.xlsx") %>% 
  select(IMO, LAT, LON, TIMESTAMP) %>% 
  distinct()

all_23_24$TIMESTAMP <- as.Date(all_23_24$TIMESTAMP)

#foreign port coordinates
port_coords <- read_csv("data/lng_ports_updated.csv") %>% 
  clean_names()

port_coords$lat <- as.numeric(port_coords$lat)
port_coords$lon <- as.numeric(port_coords$lon)

port_coords <- port_coords %>% 
  rename("LON" = "lon",
         "LAT" = "lat")

#ship identifying numbers
imo <- read_excel("data/lng_imo_values.xlsx") %>% 
  rename("tanker" = "Complete List of US Export LNG Vessels",
         "IMO" = "IMO Numbers corresponding to vessels in column A")

#ship departure info
lng_info_2017 <- read_excel("data/lng_info_2017.xlsx") %>% 
  clean_names() %>% 
  mutate(arrival_departure_date = as.Date(arrival_departure_date)) %>% 
  rename("departure_date" = "arrival_departure_date")

lng_info_2017 <- lng_info_2017 %>% 
  left_join(imo, by="tanker") %>% 
  select(departure_date, tanker, country, IMO)

lng_info_23_24 <- read_excel("data/lng_info_23_24.xlsx") %>% 
  clean_names() %>% 
  mutate(arrival_departure_date = as.Date(arrival_departure_date)) %>% 
  rename("departure_date" = "arrival_departure_date")

lng_info_23_24 <- lng_info_23_24 %>% 
  left_join(imo, by="tanker") %>% 
  select(departure_date, tanker, country, IMO)
```

### Defining the main trip duration-calculating function

```{r}
calculate_trip_durations <- function(all_positions, trip_info, port_coords, distance_threshold = 111000) {

  # Function to find the nearest port
  find_nearest_port <- function(lat, lon, ports) {
    distances <- distm(c(lon, lat), ports[, c("LON", "LAT")], fun = distHaversine)
    nearest_port <- ports[which.min(distances), ]
    return(nearest_port)
  }

  # Initialize a data frame to store trip durations
  trip_durations <- data.frame(IMO = character(), departure_date = character(), arrival_date = character(), duration_days = numeric())

  # Loop through each trip in trip_info
  for (i in 1:nrow(trip_info)) {
    trip <- trip_info[i, ]
    
    # Get departure date and IMO
    departure_date <- ymd(trip$departure_date)
    IMO <- trip$IMO
    destination_country <- trip$country
    
    # Filter positional data for the ship and after the departure date
    ship_positions <- all_positions %>% filter(IMO == IMO & TIMESTAMP >= departure_date)
    
    # Check if ship_positions is empty
    if (nrow(ship_positions) == 0) {
      next
    }
    
    # Get the coordinates of the destination port
    destination_ports <- port_coords %>% filter(country == destination_country)
    
    # Check if destination_ports is empty
    if (nrow(destination_ports) == 0) {
      next
    }
    
    # Find the arrival date
    arrival_date <- NA
    for (j in 1:nrow(ship_positions)) {
      position <- ship_positions[j, ]
      nearest_port <- find_nearest_port(position$LAT, position$LON, destination_ports)
      
      # Ensure nearest_port has the correct columns
      if (!all(c("LON", "LAT") %in% names(nearest_port))) {
        next
      }
      
      distance_to_port <- distHaversine(c(position$LON, position$LAT), c(nearest_port$LON, nearest_port$LAT))
      
      # Assume the ship arrives if it is within the specified distance threshold
      if (distance_to_port < distance_threshold) {
        arrival_date <- position$TIMESTAMP
        break
      }
    }
    
    # Calculate trip duration
    if (!is.na(arrival_date)) {
      duration_days <- as.numeric(difftime(arrival_date, departure_date, units = "days"))
      trip_durations <- rbind(trip_durations, data.frame(IMO = IMO, departure_date = departure_date, arrival_date = arrival_date, duration_days = duration_days))
    }
  }
  
  # Exclude trips that were zero days long and calculate the average trip duration
  non_zero_trip_durations <- trip_durations %>% filter(duration_days > 0)
  average_trip_duration <- mean(non_zero_trip_durations$duration_days, na.rm = TRUE)
  
  return(list(trip_durations = trip_durations, average_trip_duration = average_trip_duration))
}

```

## Comparison of overall 2017 and 2023-24 averages

```{r}
all_2017_trip_durations <- calculate_trip_durations(all_2017, lng_info_2017, port_coords, distance_threshold = 111000)
all_23_24_trip_durations <- calculate_trip_durations(all_23_24, lng_info_23_24, port_coords, distance_threshold = 111000)

cat("On average, a trip in 2017 took", all_2017_trip_durations$average_trip_duration, "days, and a trip between April 2023 and March 2024 took", all_23_24_trip_durations$average_trip_duration, "days.")
```

## Comparison by Continent

LNG vessels appear to have traveled to more countries in Asia, Europe, and North/South America in 2023 and 2024 compared to 2017. Here is the list of countries:

Asia (2017)
Japan, Jordan, India, China, Turkey, Kuwait, South Korea, Pakistan, Thailand, United Arab Emirates, Taiwan, Egypt (included as part of Asia because it was the only technically African country and it is often geographically associated with Asia) 
Asia (2023-24)
Kuwait, South Korea, India, Japan, Turkey, Taiwan, China, Thailand, Bangladesh, Singapore, Jordan, Indonesia, Philippines, Pakistan

Europe (2017)
Spain, Malta, Portugal, Netherlands, Poland, Italy, United Kingdom, Lithuania 
Europe (2023-24)
Netherlands, Belgium, France, Croatia, Spain, United Kingdom, Poland, Germany, Greece, Portugal, Italy, Lithuania, Finland

Americas (2017)
Mexico, Chile, Dominican Republic, Argentina, Brazil 
Americas (2023-24)
Dominican Republic, Brazil, Argentina, Chile, Jamaica, Colombia, Panama, Mexico, El Salvador

### Asia

```{r}
asia_2017 <- lng_info_2017 %>% 
  filter(country == "Japan"|country == "Jordan"|country == "India"|country == "China"|country == "Turkiye"|country == "Kuwait"|country == "South Korea"|country == "Pakistan"|country == "Thailand"|country == "United Arab Emirates"|country == "Taiwan"|country == "Egypt")

asia_23_24 <- lng_info_23_24 %>% 
  filter(country == "Kuwait"|country == "South Korea"|country == "India"|country == "Japan"|country == "Turkiye"|country == "Taiwan"|country == "China"|country == "Thailand"|country == "Bangladesh"|country == "Singapore"|country == "Jordan"|country == "Indonesia"|country == "Philippines"|country == "Pakistan")

asia_2017_trip_durations <- calculate_trip_durations(all_2017, asia_2017, port_coords, distance_threshold = 111000)
asia_23_24_trip_durations <- calculate_trip_durations(all_23_24, asia_23_24, port_coords, distance_threshold = 111000)

cat("For trips to Asia, the average length in 2017 was", asia_2017_trip_durations$average_trip_duration, "days, and the average length between April 2023 and March 2024 was", asia_23_24_trip_durations$average_trip_duration, "days.")
```

### Europe

```{r}
europe_2017 <- lng_info_2017 %>% 
  filter(country == "Spain"|country == "Malta"|country == "Portugal"|country == "Netherlands"|country == "Poland"|country == "Italy"|country == "United Kingdom"|country == "Lithuania")

europe_23_24 <- lng_info_23_24 %>% 
  filter(country == "Netherlands"|country == "Belgium"|country == "France"|country == "Croatia"|country == "Spain"|country == "United Kingdom"|country == "Poland"|country == "Germany"|country == "Greece"|country == "Portugal"|country == "Italy"|country == "Lithuania"|country == "Finland")

europe_2017_trip_durations <- calculate_trip_durations(all_2017, europe_2017, port_coords, distance_threshold = 111000)
europe_23_24_trip_durations <- calculate_trip_durations(all_23_24, europe_23_24, port_coords, distance_threshold = 111000)

cat("For trips to Europe, the average length in 2017 was", europe_2017_trip_durations$average_trip_duration, "days, and the average length between April 2023 and March 2024 was", europe_23_24_trip_durations$average_trip_duration, "days.")
```

### North and South America

```{r}
americas_2017 <- lng_info_2017 %>% 
  filter(country == "Mexico"|country == "Chile"|country == "Dominican Republic"|country == "Argentina"|country == "Brazil")

americas_23_24 <- lng_info_23_24 %>% 
  filter(country == "Dominican Republic"|country == "Brazil"|country == "Argentina"|country == "Chile"|country == "Jamaica"|country == "Colombia"|country == "Panama"|country == "Mexico"|country == "El Salvador")

americas_2017_trip_durations <- calculate_trip_durations(all_2017, americas_2017, port_coords, distance_threshold = 111000)
americas_23_24_trip_durations <- calculate_trip_durations(all_23_24, americas_23_24, port_coords, distance_threshold = 111000)

cat("For trips to North and South America, the average length in 2017 was", americas_2017_trip_durations$average_trip_duration, "days, and the average length between April 2023 and March 2024 was", americas_23_24_trip_durations$average_trip_duration, "days.")
```

## The Longest Trips

```{r}
trips_2017 <- as.data.frame(all_2017_trip_durations$trip_durations)

top_five_2017 <- trips_2017 %>% 
  arrange(desc(duration_days)) %>% 
  head(10)

top_five_2017 %>% 
  gt() %>% 
  cols_label(departure_date = "Departure from US",
             arrival_date = "Arrival at Foreign Port",
             duration_days = "Length of Trip (Days)") %>% 
  tab_header(title = "Top 10 Longest Trips of 2017") %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "title")) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels())
```

```{r}
trips_23_24 <- as.data.frame(all_23_24_trip_durations$trip_durations)

top_five_23_24 <- trips_23_24 %>% 
  arrange(desc(duration_days)) %>% 
  head(10)

top_five_23_24 %>% 
  gt() %>% 
  cols_label(departure_date = "Departure from US",
             arrival_date = "Arrival at Foreign Port",
             duration_days = "Length of Trip (Days)") %>% 
  tab_header(title = "Top 10 Longest Trips from April 2023 to March 2024") %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "title")) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels())
```
